---
title: "NEXP1_flower_traits_2022"
author: "Janelle Bohey"
date: "2023-10-12"
output: 
    html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
setwd("~/Desktop/ipomopsis-nitrogen")

library(car)
library(vegan)
library(dplyr)
library(emmeans)
library(lmerTest) 
library(ggplot2) 
library(tidyr)
library(tidyverse)#ggplot
library(vegan) 
library(ggpubr)
library(ggrepel)
library(googlesheets4)
library(knitr)
library(broom)
library(broom.mixed)

#set contrast 
#in order to run type 3 anova must set contrast before lmer
options(contrasts = c("contr.sum", "contr.poly")) 

#maybe for cap: options(contrasts = c("contr.treatment", "contr.poly"))for when you plot arrows


#library(multcomp) #reads lsmeans cld don't load just ::
```

# Greenhouse - read in morphology metadata & find plant averags for morph traits
```{r}

#TODO: inflo height - only have 1 measurement per plant (delete replicates)
exp1.treatments.greenhouse <-read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Treatments") 

morphdata <- read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Flower Morphology", na=c("","NA")) %>% 
  mutate(corolla_length=as.numeric(corolla_length))%>% left_join(exp1.treatments.greenhouse) %>% 
  mutate(S.T=paste0(species,treatment))

#Making new vector with means of each trait for each individual plant
morphdata1<- morphdata %>%
    group_by(plantid, treatment, species) %>% #takes mean for each plant (i.e. agg 1)
  summarise(across(c(corolla_length, corolla_width, style_length, max_anther, min_anther,sepal_width, inflo_height_cm),
              ~mean(.x, na.rm=TRUE))) 
  

```

#Field - read in morph data & find averages
```{r}
exp1.treatments.field <-read_sheet("1-da-LVmcsivkvo8MHfHbBa_LnMxGA64Ob9qCtySYANA", sheet="2022") %>% mutate(treatment=dplyr::recode(treatment, "water control"="watercontrol"))
 


morphdata.field <- read_sheet("1NNWmlf8Ts8JXoCbynnFB3l56fnt6tTup4X3ffrzN71E", sheet="2022") %>% 
   select(-treatment) %>% left_join(exp1.treatments.field) %>% mutate(treatment=replace_na(treatment, "ambient")) %>%  # get rid of this last part when you get real treatments for all plants from arturo
     mutate(treatment2=dplyr::recode(treatment, "watercontrol"= "ambient")) %>% #combines control & water control
    mutate(treatment3=dplyr::recode(treatment, "watercontrol"="water", "nitrogen"="water")) %>%
    drop_na(corolla_length) %>%  #drops plants with no data based on nothing recorded for corolla length
    mutate(plantid=paste0(site, plant)) %>% 
    mutate(S.T=paste0(site,treatment)) %>% 
    mutate(S.T2=paste0(site,treatment2)) %>% 
    mutate(S.T3=paste0(site,treatment3))
  
#Making new vector with means of each trait for each individual plant
morphdata1.field<- morphdata.field %>%
    group_by(plantid, treatment, site) %>% #takes mean for each plant (i.e. agg 1)
  summarise(across(c(corolla_length, corolla_width, style_length, max_anther, min_anther,sepal_width),
              ~mean(.x, na.rm=TRUE))) 
  
```


#Greenhouse - Model for All the traits
```{r}
library(broom)
all.morph.traits <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width", "inflo_height_cm")

morph.models <- morphdata %>% pivot_longer(all_of(all.morph.traits), names_to = "trait") %>% group_by(trait) %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~S.T + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht= map(morph.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c("AGGLOW - AGGHIGH == 0",
                                             "TENLOW - TENHIGH ==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      #formatting for kable below
      dplyr::select(-data,-morph.trait.lmer, -morph.Anova) %>% 
      unnest(morph.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )

kable(morph.models, digits=3) #spits out model into table

#boxplot for averages of all traits
allmorphtraits <- morphdata1 %>% pivot_longer(all_of(all.morph.traits), names_to = "trait") %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color= str_to_title(treatment), x=str_to_title(species), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width", inflo_height_cm="Inforecence Height (cm)")))+
    labs(y="Length (mm)", x="Nitrogen", color="Nitrogen") +
  scale_color_manual(values = c("orange","purple"))+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom")
  #scale_y_continuous(limits=c(0, NA))

print(allmorphtraits)
ggsave(allmorphtraits, file="plots-nitrogen/allmorphtraitn.png", width= 4.5, height=5.5)

#glht
multcomp:::summary.glht(multcomp::glht(total.emission.model2, linfct=multcomp::mcp(S.T= 
                                           c("agglow - agghigh == 0",
                                             "tenlow-tenhigh==0"))),                                                               test=multcomp::adjusted(type="fdr"))



```


#Field - Model for All the traits - nitrogen, water control, ambient - USE THIS ONE! 
```{r}
library(broom)
all.morph.traits.field <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width")

morph.models.field <- morphdata.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% group_by(trait)          %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~S.T + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht= map(morph.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c("Laggambient - Laggnitrogen == 0",
                                             "Ihybambient - Ihybnitrogen == 0",
                                             "Laggwatercontrol - Laggambient == 0",
                                             "Ihybwatercontrol - Ihybambient == 0",
                                             "Laggnitrogen - Laggwatercontrol==0",
                                             "Ihybnitrogen - Ihybwatercontrol==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      #formatting for kable below
      dplyr::select(-data,-morph.trait.lmer, -morph.Anova) %>% 
      unnest(morph.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )

kable(morph.models.field, digits=3) #spits out model into table

#boxplot for averages of all traits
allmorphtraits.field <- morphdata1.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color= str_to_title(treatment), x=str_to_title(site), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width", inflo_height_cm="Inforecence Height (cm)")))+
    labs(y="Length (mm)", x="Nitrogen", color="Nitrogen") +
  scale_color_manual(values = c("orange","purple","blue"))+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom")
  #scale_y_continuous(limits=c(0, NA))

print(allmorphtraits.field)

ggsave(allmorphtraits.field, file="plots-nitrogen/NEXP1.field.allmorphtraits.png", width= 4.5, height=5.5)

#difference in corolla width is due to watering plant not due to nitrogen 

```

#Field - Model for All the traits - ONLY nitrogen, ambient = water effect
```{r}
library(broom)
all.morph.traits.field <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width")

morph.models.field <- morphdata.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% group_by(trait)          %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~S.T2 + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht= map(morph.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T2= 
                                           c("Laggambient - Laggnitrogen == 0",
                                             "Ihybambient - Ihybnitrogen == 0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      #formatting for kable below
      dplyr::select(-data,-morph.trait.lmer, -morph.Anova) %>% 
      unnest(morph.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )

kable(morph.models.field, digits=3) #spits out model into table

#boxplot for averages of all traits
allmorphtraits.field <- morphdata1.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color= str_to_title(treatment), x=str_to_title(site), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width", inflo_height_cm="Inforecence Height (cm)")))+
    labs(y="Length (mm)", x="Nitrogen", color="Nitrogen") +
  scale_color_manual(values = c("orange","purple","blue"))+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom")
  #scale_y_continuous(limits=c(0, NA))

print(allmorphtraits.field)

ggsave(allmorphtraits.field, file="plots-nitrogen/NEXP1.field.allmorphtraits.png", width= 4.5, height=5.5)

#difference in corolla width is due to watering plant not due to nitrogen 

```

#Field - Model for All the traits - ONLY water =(nitrogen+water control), ambient 
```{r}
library(broom)
all.morph.traits.field <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width")

morph.models.field <- morphdata.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% group_by(trait)          %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~S.T3 + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht= map(morph.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T3= 
                                           c("Laggambient - Laggwater == 0",
                                             "Ihybambient - Ihybwater == 0" 
                                              ))), test=multcomp::adjusted(type="fdr"))))) %>% 
      #formatting for kable below
      dplyr::select(-data,-morph.trait.lmer, -morph.Anova) %>% 
      unnest(morph.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )

kable(morph.models.field, digits=3) #spits out model into table

#boxplot for averages of all traits
allmorphtraits.field <- morphdata1.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color= str_to_title(treatment), x=str_to_title(site), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width", inflo_height_cm="Inforecence Height (cm)")))+
    labs(y="Length (mm)", x="Nitrogen", color="Nitrogen") +
  scale_color_manual(values = c("orange","purple","blue"))+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom")
  #scale_y_continuous(limits=c(0, NA))

print(allmorphtraits.field)

ggsave(allmorphtraits.field, file="plots-nitrogen/NEXP1.field.allmorphtraits.png", width= 4.5, height=5.5)

#difference in corolla width is due to watering plant not due to nitrogen 

```

# Greenhouse - nectar
```{r}
exp1.treatments <-read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Treatments")

nectardata <- read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Nectar") %>% 
  #mutate(plant = as.numeric(plant)) %>% 
  mutate(nectar.production = nectar_48_h_mm * 5 /(2 * 32)) %>%    #5-uL microcapillary tube 32 mm in length, 2 days
  left_join(exp1.treatments) %>% mutate(treatment=factor(treatment, levels=c("LOW","HIGH")))

nectar.traits <- c('nectar.production', "concentration") 

library(broom)

nectar.models <- nectardata %>% pivot_longer(all_of(nectar.traits), names_to = "trait") %>% group_by(trait,species) %>% nest() %>% 
      mutate(nectar.trait.lmer=map(data, ~lmer(value~treatment+ (1|plantid), data=.x))) %>% 
      mutate(nectar.Anova= map(nectar.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  unnest(nectar.Anova) %>%
      dplyr::select(-data,-nectar.trait.lmer) %>% 
      mutate(significant = ifelse(p.value<0.05, "*","") ) %>% filter(term!="(Intercept)")

kable(nectar.models, digits=3) #spits out model into table

#average nectar data
 nectardata1<- nectardata %>%
    group_by(plantid, treatment, species, plant_num) %>% 
    summarise(across(all_of(nectar.traits), ~mean(.x, na.rm=TRUE))) 

#boxplot for averages of all traits
 
nectardata1 %>% pivot_longer(all_of(nectar.traits), names_to = "trait") %>% #need to pivot longer for facets
    ggplot(aes(color= treatment, x=species, y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y")+
    ylab("") +
   scale_color_manual(values = c("purple", "orange"))

#
nectardata %>% ggplot(aes(x=concentration  , y=nectar.production))+geom_point()+geom_smooth()

```

# Field - nectar 
```{r}
exp1.treatments.field <-read_sheet("1-da-LVmcsivkvo8MHfHbBa_LnMxGA64Ob9qCtySYANA", sheet="2022")

nectardata.field <- read_sheet("1NNWmlf8Ts8JXoCbynnFB3l56fnt6tTup4X3ffrzN71E", sheet="nectar", na=c("","NA")) %>% 
    drop_na(nectar_48_h_mm) %>% 
  #mutate(plant = as.numeric(plant)) %>% 
  mutate(nectar.production = nectar_48_h_mm * 5 /(2 * 32)) %>%    #5-uL microcapillary tube 32 mm in length, 2 days
  left_join(exp1.treatments.field) %>% mutate(plantid=paste0(site,plant)) %>% mutate(treatment=replace_na(treatment, "ambient"))

nectar.traits.field <- c('nectar.production', "nectar_conc") 

nectar.field.models <- nectardata.field %>% pivot_longer(all_of(nectar.traits.field), names_to = "trait") %>% group_by(trait,site) %>% nest() %>% 
      mutate(nectar.trait.field.lmer=map(data, ~lmer(value~treatment+ (1|plantid), data=.x))) %>% 
      mutate(nectar.field.Anova= map(nectar.trait.field.lmer, ~tidy(Anova(.x, type=3)))) %>%  unnest(nectar.field.Anova) %>%
      dplyr::select(-data,-nectar.trait.field.lmer) %>% 
      mutate(significant = ifelse(p.value<0.05, "*","") ) %>% filter(term!="(Intercept)")

kable(nectar.field.models, digits=3) #spits out model into table

#average nectar data
 nectardata1.field<- nectardata.field %>%
    group_by(plantid, treatment, site, plant) %>% 
    summarise(across(all_of(nectar.traits.field), ~mean(.x, na.rm=TRUE))) 

#boxplot for averages of all traits
 
nectardata1.field %>% pivot_longer(all_of(nectar.traits.field), names_to = "trait") %>% #need to pivot longer for facets
    ggplot(aes(color= treatment, x=site, y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y")+
    ylab("") +
   scale_color_manual(values = c("purple", "orange", "blue"))

#
nectardata.field %>% ggplot(aes(x=nectar.production  , y=nectar_conc))+geom_point()+geom_smooth()

```


#Total Flower number
```{r}
total.flower<-read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Flower Number") %>% 
   left_join(exp1.treatments) %>% mutate(treatment=factor(treatment, levels=c("LOW","HIGH")))


flower.nub.lm=lm(flower_num~treatment*species, data=total.flower)
  
Anova(flower.nub.lm, type=3)#Type 3

ggplot(total.flower, aes(y=flower_num, x=species, color=treatment))+geom_boxplot()+
  scale_color_manual(values = c("purple", "orange"))

#nitrogen only p= 0.0074

#######old way#############################
#make data frame with ls means
flower.num.N= as.data.frame(lsmeans(flower.nub.lmer, pairwise~ nitrogen|microbe, adjustSigma = TRUE, adjust = "tukey")$lsmeans)

#contrast of treatment (high vs low N) in each species (TEN vs AGG), groups a, b
multcomp::cld(lsmeans(flower.nub.lmer, pairwise~ nitrogen|microbe, adjustSigma = TRUE, adjust = "tukey")$lsmeans ,
alpha=0.05, Letters=letters,adjust="tukey")

###### glht #########

#library (tibble)

#need treatmentspecies so paste0 
flowerdatamutate <- as_tibble(total.flower)%>%mutate(nitrogenmicrobe=paste0(nitrogen,microbe))%>% 
  mutate(nitrogenmicrobe=factor(nitrogenmicrobe))#paste0 because glht doesn't like spaces in data sets

#model to put into glht bc glht doesn't like spaces
flower.nub.lmer.glht=lmer(total_flower_number~nitrogenmicrobe + (1|plant), data=flowerdatamutate)

summary(multcomp::glht(flower.nub.lmer.glht, linfct=multcomp::mcp(nitrogenmicrobe="Tukey")), test=multcomp::adjusted(type="none"))

##plot with species on x-axis, bars show N treatment, y axis is lsmean

Fig.flower.number = ggplot(flower.num.N, aes(x=nitrogen, y=lsmean, fill=microbe)) +
  geom_bar(position=position_dodge(), stat="identity", width = 0.5,  linewidth=0.5)+
  geom_errorbar(aes(ymin=lsmean-SE, ymax=lsmean+SE),
                width=.2, # Width of the error bars
                position=position_dodge(.5))  +
scale_fill_manual(values=c("orange", "lightblue3"), name = "Nitrogen Level", labels = c("High", "Low")) + 
  labs(y="Total Flower Number", x="") +
  annotate("text", x = 1.6, y = 300, label = "*", size = 4)+
  annotate("text", x = 1, y = 280, label = "**", size = 4)+
  theme(panel.border = element_blank(),
      panel.grid.major = element_blank(),
      #panel.grid.minor = element_blank(),
      panel.background = element_blank())+
  theme(legend.title = element_text(size=11, color = "black"),
        legend.text=element_text(size=9),
        #axis.text.x = element_text(size=9),
        axis.ticks = element_blank())#+
  #scale_x_discrete(labels=c('I. aggregata', 'I. tenuituba'))

Fig.flower.number
ggsave(Fig.flower.number, file='plot_flower_number.jpg', width=110, height=110, units= 'mm', dpi=300)


```

