---
title: "NEXP1_flower_traits_2022"
author: "Janelle Bohey"
date: "2023-10-12"
output: 
    html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
setwd("~/Desktop/ipomopsis-nitrogen")

library(car)
library(vegan)
library(dplyr)
library(emmeans)
library(lmerTest) 
library(ggplot2) 
library(tidyr)
library(tidyverse)#ggplot
library(vegan) 
library(ggpubr)
library(ggrepel)
library(googlesheets4)
library(knitr)
library(broom)
library(broom.mixed)
   # install.packages("ggpubr")
    library(ggpubr)

#set contrast 
#in order to run type 3 anova must set contrast before lmer
options(contrasts = c("contr.sum", "contr.poly")) 

#maybe for cap: options(contrasts = c("contr.treatment", "contr.poly"))for when you plot arrows

```

#data visualization
```{r,data visualization set up}
#set standard figure dimensions
opts_chunk$set(fig.height=8, fig.width=8, dpi=100, fig.path="figures/") #sets standard width and height for all figures in rmd; creates folder only when rmd is knitted
#can also change height in {} at the top of the chunk

#set figure theme
NEXP1.theme <- theme(
      #axis text
                  axis.title = element_text(size=13),     #x axis title text size
                  axis.text.x = element_text(size=12, color = "black"),     #x axis labels text size
                  axis.ticks.x = element_blank(),  # Remove tick marks on x-axis
                  axis.title.x = element_text(margin = margin(t = 10)),  # Adds 10 units of space on top of the x-axis title
                  axis.title.y = element_text(margin = margin(r = 15)),  # Adds 15 units of space to the right of the y-axis title
      #legend
                  legend.text = element_text(size=11),       #legend text size
                  legend.title = element_text(size=12, color = "black", hjust = 0.5),  #legend title text size
                  legend.position = "right",                #"left", "right", "bottom", "top", or a coordinate
                  legend.direction = "vertical",           #layout direction of the legend ("horizontal" or "vertical")
                  legend.spacing = unit(1, "cm"),            #spacing between legend items; adds 1cm of space between
                  legend.key.size = unit(1.1, "cm"),         #size of the legend keys (symbols); sets legend key to 1.5cm
      #panels
                  panel.border = element_blank(),            # border around the plot area
                  panel.grid.major = element_line(colour="white"),
                  #panel.grid.minor = element_blank(),
                  panel.background = element_rect(fill = "white", # # Remove grey background (set to white)
                                          #colour = "lightblue",
                                          linetype = "solid"),
      
      #facets
                  strip.text.x = element_text(size=14),       #strip text on x-axis facet
                  strip.background = element_rect(colour="white", fill="white")) #background for the strip label
                  strip.text = element_text(size = 16)        #This changes the size of the facet labels


field.label.colors.theme <- scale_color_manual(name="Nitrogen Level", 
          labels= c(ambient="No water added", nitrogen= "High nitrogen", watercontrol= "Low nitrogen"), 
          values=c(ambient="#457A26",nitrogen="#E69F00", watercontrol="#0072B2"))

greenhouse.label.colors.theme <- scale_color_manual(name="Nitrogen Level", 
          labels= c( HIGH= "High nitrogen", LOW= "Low nitrogen"), 
          values=c(HIGH="#E69F00", LOW="#0072B2"))

#color scheme
#scale_color_manual(name="Nitrogen Level", 
#        labels= c("no water added",  "high nitrogen", "low nitrogen"), #Specifies treatment names in legend
#        values=c("#8b9b31", "#b35f98", "#3a8db3")
  

#ggplot boxplot code: 
#plot.total.emissions.field=
#  ggplot(meta.field, aes( y= totalemission, x=time, color=treatment))+
#  geom_boxplot(position=position_dodge(width=0.9))+
#  geom_point (position=position_dodge(width=0.9))+ #makes it so point don't overlap
#  scale_y_sqrt()+
#  ylab("Total Emissions")+
#  xlab("") +
#  facet_wrap(vars(species)) +
#  facet_grid(. ~ species, labeller=labeller(species = facet.labs))+ 
#  scale_x_discrete(labels=c('Day', 'Night', 'Day','Night'))+ #x axis text
#  ylim(0.1, 0.56)+  # Set y-axis limits from 0.1 to 0.56 + 
#  scale_color_manual(name="Nitrogen Level", 
#        labels= c("no water added",  "high nitrogen", "low nitrogen"), #Specifies treatment names in legend
#        values=c("#8b9b31", "#b35f98", "#3a8db3")


#print(plot.total.emissions.field)  
```


# Greenhouse - read in morphology metadata & find plant averags for morph traits
```{r}

exp1.treatments.greenhouse <-read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Treatments") #loads from EXP 1 (N) Flower Traits on google drive

morphdata <- read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Flower Morphology", na=c("","NA")) %>% 
  mutate(corolla_length=as.numeric(corolla_length))%>% left_join(exp1.treatments.greenhouse) %>% 
  mutate(S.T=paste0(species,treatment)) %>% 
  #set treatment levels
      mutate(treatment = as.character(treatment)) %>% 
  mutate(treatment = factor(treatment, levels = c("LOW", "HIGH")))


#Making new vector with means of each trait for each individual plant
morphdata1<- morphdata %>%
    group_by(plantid, treatment, species) %>% #takes mean for each plant (i.e. agg 1)
  summarise(across(c(corolla_length, corolla_width, style_length, max_anther, min_anther,sepal_width, inflo_height_cm),
              ~mean(.x, na.rm=TRUE))) 
  
  

```

#Field - read in morph data & find averages
```{r}
exp1.treatments.field <-read_sheet("1-da-LVmcsivkvo8MHfHbBa_LnMxGA64Ob9qCtySYANA", sheet="2022") %>% mutate(treatment=dplyr::recode(treatment, "water control"="watercontrol"))
 


morphdata.field <- read_sheet("1NNWmlf8Ts8JXoCbynnFB3l56fnt6tTup4X3ffrzN71E", sheet="2022") %>% 
   dplyr::select(-treatment) %>% left_join(exp1.treatments.field) %>% mutate(treatment=replace_na(treatment, "ambient")) %>%  # get rid of this last part when you get real treatments for all plants from arturo
     mutate(treatment2=dplyr::recode(treatment, "watercontrol"= "ambient")) %>% #combines control & water control
    mutate(treatment3=dplyr::recode(treatment, "watercontrol"="water", "nitrogen"="water")) %>% 
   mutate(site=factor(site, levels=c("Lagg","Ihyb"))) %>% 
  #set treatment levels for figures
  mutate(treatment = factor(treatment, levels = c("ambient", "watercontrol", "nitrogen"))) %>% 
    drop_na(corolla_length) %>%  #drops plants with no data based on nothing recorded for corolla length
    mutate(plantid=paste0(site, plant)) %>% 
    mutate(S.T=paste0(site,treatment)) %>% 
    mutate(S.T2=paste0(site,treatment2)) %>% 
    mutate(S.T3=paste0(site,treatment3)) 
  
#Making new vector with means of each trait for each individual plant
morphdata1.field<- morphdata.field %>%
    group_by(plantid, treatment, site) %>% #takes mean for each plant (i.e. agg 1)
  summarise(across(c(corolla_length, corolla_width, style_length, max_anther, min_anther,sepal_width),
              ~mean(.x, na.rm=TRUE))) 
  
```

# Save morph data for VOC correlation in ipo_nitrogen.rmd
```{r}
write_csv(morphdata1, "data/morphdata1.csv")

write_csv(morphdata1.field, "data/morphdata1.field.csv")

```

#MORPH 
##GH- Model for All the traits
```{r}
library(broom)
library(emmeans)
all.morph.traits <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width")


morph.models.everything <- morphdata %>% pivot_longer(all_of(all.morph.traits), names_to = "trait") %>%
  group_by(trait) %>%    
  nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~S.T + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht= map(morph.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c("AGGLOW - AGGHIGH == 0",
                                             "TENLOW - TENHIGH ==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      mutate(emmeans=map(morph.trait.lmer, ~tidy(emmeans(.x, "S.T")))) #gives means so we can do % change from high to low

      #formatting for kable below
morph.models<- morph.models.everything %>% dplyr::select(-data,-morph.trait.lmer, -morph.Anova, -emmeans) %>% 
      unnest(morph.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )  
      

kable(morph.models, digits=3) #spits out model into table

#emmeans table 
morph.emmeans<- morph.models.everything %>% dplyr::select(-data,-morph.trait.lmer, -morph.Anova, -morph.glht) %>% 
      unnest(emmeans)  
    
kable(morph.emmeans, digits=3) #spits out model into table
#now divide high by low to get % change




#boxplot for averages of all traits

allmorphtraits <- morphdata1 %>% filter(plantid != "TEN1") %>% #removes TEN1 bc no data except inflo height
          pivot_longer(all_of(all.morph.traits), names_to = "trait") %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color= treatment, x=str_to_title(species), y=value))+
        geom_boxplot(position=position_dodge(width=0.9))+
        geom_point (position=position_dodge(width=0.9))+ #makes it so point don't overlap
        facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width")))+
        labs(y="Length (mm)", x= "", color="Nitrogen") +
   #scale_y_sqrt()+
   scale_color_manual(name="Nitrogen Level", 
          labels= c( HIGH= "water + nitrogen", LOW= "water only"), #Lower case bc str_to_title chagnes treatment name
          values=c(HIGH="#E69F00", LOW="#0072B2")) +
  NEXP1.theme + 
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80", size = 0.5),
        legend.position = "right")+
  scale_y_continuous(limits = c(0, NA), expand=expansion(mult=c(0.02,0.2))) +
  scale_x_discrete(
    labels = c(
      "Ten" = expression(italic("I. tenuituba")),
      "Agg" = expression(italic("I. aggregata"))
    )) +  
  theme(
    axis.text.x = element_text(size = 11), # bigger x-axis text
    legend.text = element_text(size = 10) 
  )


print(allmorphtraits)

ggsave(allmorphtraits, file="plots-nitrogen/allmorphtraits.png", width=220, height=180, units= 'mm', dpi=600)


```


##Field - Model for All the traits - nitrogen, water control, ambient - USE THIS ONE! 
```{r}
library(broom)
all.morph.traits.field <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width")

morph.models.field <- morphdata.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% group_by(trait) %>% nest() %>% 
      mutate(morph.trait.lmer.field=map(data, ~lmer(value~S.T + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova.field= map(morph.trait.lmer.field, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht.field= map(morph.trait.lmer.field, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c(
                                             "Laggwatercontrol - Laggambient == 0",
                                             "Ihybwatercontrol - Ihybambient == 0",
                                             "Laggnitrogen - Laggwatercontrol==0",
                                             "Ihybnitrogen - Ihybwatercontrol==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
    mutate(morph.emmeans.field=map(morph.trait.lmer.field, ~tidy(emmeans(.x, "S.T")))) #gives means so we can do % change from high to low


      #formatting for kable below
morph.models.glht.field<- morph.models.field %>% dplyr::select(-data,-morph.trait.lmer.field, -morph.Anova.field, -morph.emmeans.field) %>% 
      unnest(morph.glht.field) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )  
      

kable(morph.models.glht.field, digits=3) #spits out model into table

#emmeans table 
morph.emmeans.field<- morph.models.field %>% dplyr::select(-data,-morph.trait.lmer.field, -morph.Anova.field, -morph.glht.field) %>% 
      unnest(morph.emmeans.field)  
    
kable(morph.emmeans.field, digits=3)


######
# Create a tibble with bracket coordinates and labels for corolla_width
#HYBRIDS ON LEFT, Lagg ON RIGHT
brackets_corolla_width_Hybrids_Lagg <- tibble(
  trait = c("corolla_width", "corolla_width"),  # only for corolla_width
  right = c(0.83, 2.2),  # Move the right bracket further to the right
  left = c(0.79, 2.1),   # Shorten the left bracket distance
  label = c("p = 0.015*", "p = 0.024*"),  # p-values or other annotations
  y.position = c(4, 4.5)  # Adjusted y position for the brackets
)

#Lagg ON LEFT, HYBRIDS on right
######
# Create a tibble with bracket coordinates and labels for corolla_width
brackets_corolla_width_Lagg_Hybrids <- tibble(
  trait = c("corolla_width", "corolla_width"),  # only for corolla_width
  right = c(1.84, 0.5),    # Adjust the right bracket's position
  left  = c(1.8, 1.25),    # Slightly adjust left bracket to be closer to Lagg (right side)
  label = c("p = 0.024*", "p = 0.015*"),
  y.position = c(3.8, 4.6)
)

#boxplot for averages of all traits
allmorphtraits.field <- morphdata1.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% #mutate(site=factor(site, levels=c("Lagg","Ihyb"))) %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color=treatment, x=site, y=value))+
       geom_boxplot(position=position_dodge(width=0.9))+
       geom_point (position=position_dodge(width=0.9))+ #makes it so point don't overlap
       facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width")))+
       labs(y="Length (mm)", x="", color="Nitrogen") +
       NEXP1.theme+
       
       scale_color_manual(name="Nitrogen Level", 
          labels= c(ambient="No water", nitrogen= "water + nitrogen", watercontrol= "water only"), 
          values=c(ambient="#457A26", watercontrol="#0072B2",nitrogen="#E69F00")) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80", size = 0.5),
        legend.position = "right")+
       scale_y_continuous(limits = c(0, NA), expand=expansion(mult=c(0.02,0.2))) +
   geom_bracket(
  data = brackets_corolla_width_Hybrids_Lagg, 
  aes(xmin = right, xmax = left, label = label, y.position = y.position, group = interaction(trait, label)), 
  size = 0.5, bracket.shorten = 0.4, label.size = 3.5, inherit.aes = FALSE)+ 
    scale_x_discrete(
    labels = c(
      "Ihyb" = "Hybrid",
      "Lagg" = expression(italic("I. aggregata"))
    )) +  
  theme(
    axis.text.x = element_text(size = 11), # bigger x-axis text
    legend.text = element_text(size = 10) 
  )



print(allmorphtraits.field)

ggsave(allmorphtraits.field, file="plots-nitrogen/NEXP1.field.allmorphtraits.png", width=220, height=180, units= 'mm', dpi=600)

#difference in corolla width is due to watering plant not due to nitrogen 

```

#NECTAR
## GH - nectar
```{r}
exp1.treatments <-read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Treatments")

nectardata <- read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Nectar") %>%
     #mutate(plant = as.numeric(plant)) %>% 
   #5-uL microcapillary tube 32 mm in length, 2 days
     mutate(nectar.production = nectar_48_h_mm * 5 /(2 * 32)) %>%  
    left_join(exp1.treatments) %>% 
     mutate(treatment=factor(treatment, levels=c("LOW","HIGH"))) %>% 
  #create column with speciestreatment aka S.T
    mutate(S.T=paste0(species,treatment)) 

nectar.traits <- c('nectar.production', "concentration") 

#make model 
nectar.model <- nectardata %>% pivot_longer(all_of(nectar.traits), names_to = "trait") %>% group_by(trait) %>%        nest() %>% 
      mutate(nectar.trait.lmer=map(data, ~lmer(value~S.T + (1|plantid), data=.x))) %>% 
      mutate(nectar.Anova= map(nectar.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  #unnest(nectar.Anova) %>%
      #dplyr::select(-data,-nectar.trait.lmer) %>% 
      #mutate(significant = ifelse(p.value<0.05, "*","") ) %>% filter(term!="(Intercept)")

       mutate(nectar.glht= map(nectar.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c("AGGLOW - AGGHIGH == 0",
                                             "TENLOW - TENHIGH ==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      mutate(emmeans=map(nectar.trait.lmer, ~tidy(emmeans(.x, "S.T")))) #gives means so we can do % change from high to low


     #formatting glht table for kable below
nectar.glht<- nectar.model %>% dplyr::select(-data,-nectar.trait.lmer, -nectar.Anova, -emmeans) %>% 
      unnest(nectar.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )  

kable(nectar.glht, digits=3) #spits out model into table

#emmeans table 
nectar.emmeans<-nectar.model %>% dplyr::select(-data,-nectar.trait.lmer, -nectar.Anova, -nectar.glht) %>% 
      unnest(emmeans)  
    
kable(nectar.emmeans, digits=3)



#.....................................

#average nectar data
 nectardata1<- nectardata %>%
    group_by(plantid, treatment, species, plant_num) %>% 
    summarise(across(all_of(nectar.traits), ~mean(.x, na.rm=TRUE))) 


 
#boxplot for averages of all traits
  # Define new facet labels
facet_labels <- c(
  "concentration" = "Nectar Concentration (%)", #percent sucrose equivalent (degrees brix) or mass percent
  "nectar.production" = "Nectar Production (\U03BCL)"
)
facet.labs <- c(AGG="I. aggregata", TEN="I. tenuituba")

#add brackets
brackets.gh.nectar <- tibble(
  trait = c("concentration", "concentration", "nectar.production", "nectar.production"),  # 4 entries
  right = c(0.45, 1.45, 0.45, 1.45), 
  left = c(1.55, 2.55, 1.55, 2.55),
  label = c("p= 0.367", "p= 0.955", "p= 0.360", "p=0.042*"),
  y.position = c(35, 41, 6.8, 1.7)   # different y positions for each bracket
)

#plot
nectar.plot <- nectardata1 %>% pivot_longer(all_of(nectar.traits), names_to = "trait") %>% #need to pivot longer for facets
    ggplot(aes(color= treatment, x=species, y=value))+
  geom_boxplot(position=position_dodge(width=0.9))+
       geom_point (position=position_dodge(width=0.9))+ 
     facet_wrap(vars(trait), scales = "free_y", labeller = labeller(trait = facet_labels)) +
    ylab("") +
   scale_color_manual(name="Nitrogen Level", 
          labels= c( HIGH= "water + nitrogen", LOW= "water only"), 
          values=c(HIGH="#E69F00", LOW="#0072B2")) +
  scale_x_discrete(
    labels = c(
      "TEN" = expression(italic("I. tenuituba")),
      "AGG" = expression(italic("I. aggregata"))
    )
  ) +
  theme(panel.grid.major.y = element_line(color = "grey80", size = 0.5), panel.grid.major.x = element_blank(), legend.position = "bottom")+ 
   labs(y="", x="", color="Nitrogen") +
  NEXP1.theme +
    scale_y_continuous(limits = c(0, NA), expand=expansion(mult=c(0.02,0.2))) + 
 geom_bracket(
  data = brackets.gh.nectar, 
  aes(xmin = right, xmax = left, label = label, y.position = y.position, group = interaction(trait, label)), 
  size = 0.5, bracket.shorten = 0.4, label.size = 4, inherit.aes = FALSE)+ 
  theme(
    axis.text.x = element_text(size = 14,  hjust = 0.5, vjust = 0.5, 
                               margin = margin(t = 14)),        # Increase x-axis text size
    axis.text.y = element_text(size = 18),        # Increase y-axis text size
    axis.title.x = element_text(size = 20),       # Increase x-axis title size
    axis.title.y = element_text(size = 20),       # Increase y-axis title size
    legend.text = element_text(size = 18),        # Increase legend text size
    legend.title = element_text(size = 20),       # Increase legend title size
    strip.text = element_text(size = 20),         # Increase facet label size
    plot.title = element_text(size = 22),         # Increase plot title size
    plot.subtitle = element_text(size = 20)      # Increase subtitle size (if any)
  )
  

print(nectar.plot)

ggsave(nectar.plot, file='plots-nitrogen/greenhouse_nectar.jpg',  width=250, height=160, units= 'mm', dpi=600)

#nectardata %>% ggplot(aes(x=concentration  , y=nectar.production))+geom_point()+geom_smooth()

```

## Field - nectar 
```{r}
exp1.treatments.field <-read_sheet("1-da-LVmcsivkvo8MHfHbBa_LnMxGA64Ob9qCtySYANA", sheet="2022") %>% mutate(treatment=dplyr::recode(treatment, "water control"="watercontrol"))

nectardata.field <- read_sheet("1NNWmlf8Ts8JXoCbynnFB3l56fnt6tTup4X3ffrzN71E", sheet="nectar", na=c("","NA")) %>% 
    drop_na(nectar_48_h_mm) %>% 
  #mutate(plant = as.numeric(plant)) %>% 
  mutate(nectar.production = nectar_48_h_mm * 5 /(2 * 32)) %>%    #5-uL microcapillary tube 32 mm in length, 2 days
  left_join(exp1.treatments.field) %>% mutate(plantid=paste0(site,plant)) %>% mutate(treatment=replace_na(treatment, "ambient")) %>% 
  mutate(S.T=factor(paste0(site,treatment))) %>% 
  rename("concentration" = nectar_conc)

nectar.traits.field <- c('nectar.production', "concentration") 

nectar.field.model <- nectardata.field %>% pivot_longer(all_of(nectar.traits.field), names_to = "trait") %>% group_by(trait) %>% nest() %>% 
      mutate(nectar.trait.field.lmer=map(data, ~lmer(value~S.T+ (1|plantid), data=.x))) %>% 
      mutate(nectar.field.Anova= map(nectar.trait.field.lmer, ~tidy(Anova(.x, type=3)))) %>%
  #unnest(nectar.field.Anova) %>%
   #   dplyr::select(-data,-nectar.trait.field.lmer) %>% 
   #   mutate(significant = ifelse(p.value<0.05, "*","") ) %>% filter(term!="(Intercept)")
      mutate(nectar.glht.field= map(nectar.trait.field.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c(
                                             "Laggwatercontrol - Laggambient == 0",
                                             "Ihybwatercontrol - Ihybambient == 0",
                                             "Laggnitrogen - Laggwatercontrol==0",
                                             "Ihybnitrogen - Ihybwatercontrol==0"))),
      #don't do fdr bc then pvalues are all the same                                     
      test=multcomp::adjusted(type="none"))))) %>% 

      mutate(nectar.emmeans.field=map(nectar.trait.field.lmer, ~tidy(emmeans(.x, "S.T"))))
      #gives means so we can do % change from high to low


     #formatting glht table for kable below
nectar.glht.field<- nectar.field.model %>% dplyr::select(-data,-nectar.trait.field.lmer, -nectar.field.Anova, -nectar.emmeans.field) %>% 
      unnest(nectar.glht.field)  

kable(nectar.glht.field, digits=3) #spits out model into table

#emmeans table 
nectar.emmeans.field <-nectar.field.model %>% dplyr::select(-data,-nectar.trait.field.lmer, -nectar.field.Anova, -nectar.glht.field) %>% 
      unnest(nectar.emmeans.field)  
    
kable(nectar.emmeans.field, digits=3)



#average nectar data
 nectardata1.field<- nectardata.field %>%
    group_by(plantid, treatment, site, plant) %>% 
    summarise(across(all_of(nectar.traits.field), ~mean(.x, na.rm=TRUE))) 

#boxplot for averages of all traits
 # Define new facet labels
facet_labels <- c(
  "concentration" = "Nectar Concentration (%)",
  "nectar.production" = "Nectar Production (\U03BCL)"
)
facet.labs <- c(Lagg="I. aggregata", Ihyb="Hybrids")

#add brackets to ggplot
brackets.field.nectar <- tibble(
   trait = c(
    "nectar.production", "nectar.production", "nectar.production", "nectar.production",
    "concentration", "concentration", "concentration", "concentration"
  ),
    xmin = c(0.5, 0.9, 1.5, 1.9, 0.5, 0.9, 1.5, 1.9),  # wider gap between 2nd & 3rd
  xmax = c(1.0, 1.4, 2.0, 2.4, 1.0, 1.4, 2.0, 2.4), 
  y.position = c(3.5, 4.2, 5.7, 6.3, 41, 46, 34, 38),
  label = c("p=0.528", "p=0.655", "p=0.956", "p=0.022 *","p=0.753", "p=0.972", "p=0.922", "p=0.264")
)

#plot
nectar.field.plot <- nectardata1.field %>% pivot_longer(all_of(nectar.traits.field), names_to = "trait") %>% #need to pivot longer for facets
  mutate(treatment = factor(treatment, levels = c("ambient", "watercontrol", "nitrogen"))) %>% #reorders treatments to be no water, low N, high N
  ggplot(aes(color= treatment, x=site, y=value))+
  geom_boxplot(position=position_dodge(width=0.8))+
       geom_point (position=position_dodge(width=0.8))+ 
    facet_wrap(vars(trait), scales = "free_y", labeller = labeller(trait = facet_labels)) +
    ylab("") +
    scale_color_manual(name="Nitrogen Level", 
          labels= c(ambient="No water", nitrogen= "water + nitrogen", watercontrol= "water only"), 
          values=c(ambient="#457A26",nitrogen="#E69F00", watercontrol="#0072B2"))+
  scale_x_discrete(
    labels = c(
      "Ihyb" = "Hybrids",
      "Lagg" = expression(italic("I. aggregata"))
    )
  ) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80", size = 0.5),
        legend.position = "bottom")+ 
   labs(y="", x="", color="Nitrogen") +
  NEXP1.theme+
  scale_y_continuous(limits = c(0, NA), expand=expansion(mult=c(0.02,0.2))) +
   geom_bracket(
    data = brackets.field.nectar,
    aes(xmin = xmin, xmax = xmax, y.position = y.position),
    inherit.aes = FALSE,
    size = 0.5,
    bracket.shorten = 0.8,
    tip.length = 0.02
  )

print(nectar.field.plot)

ggsave(nectar.field.plot, file='plots-nitrogen/field_nectar.jpg',  width=200, height=160, units= 'mm', dpi=600)

#
nectardata.field %>% ggplot(aes(x=nectar.production  , y=nectar_conc))+geom_point()+geom_smooth()

```
#GH ONLY

##Inflo height
```{r}

inflo.model <- morphdata %>% mutate(S.T=factor(S.T)) %>%  pivot_longer(inflo_height_cm, names_to = "trait") %>% group_by(trait) %>% nest() %>% 
      mutate(inflo.lm=map(data, ~lm(value~S.T, data=.x))) %>% 
      mutate(inflo.Anova= map(inflo.lm, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(inflo.glht= map(inflo.lm, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c("AGGLOW - AGGHIGH == 0",
                                             "TENLOW - TENHIGH ==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      mutate(emmeans=map(inflo.lm, ~tidy(emmeans(.x, "S.T")))) #gives means so we can do % change from high to low

      #formatting glht table for kable below
inflo.glht<- inflo.model %>% dplyr::select(-data,-inflo.lm, -inflo.Anova, -emmeans) %>% 
      unnest(inflo.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )  

kable(inflo.glht, digits=3) #spits out model into table

#emmeans table 
inflo.emmeans<-inflo.model %>% dplyr::select(-data,-inflo.lm, -inflo.Anova, -inflo.glht) %>% 
      unnest(emmeans)  
    
kable(inflo.emmeans, digits=3) #spits out model into table
#now divide 'estimate' high by low to get % change
#AGG high/low = 48.66/40 = 1.2165 --> 21.65% increase in high????

inflo.data <- morphdata %>% select(plantid, species, treatment, inflo_height_cm) %>% drop_na()

#plot inflo height
#rename species facet labs
facet.labs <- c(AGG="I. aggregata", TEN="I. tenuituba")

#add brackets to plot
brackets.inflo <- tibble(
  right = c(0.45, 1.45), 
  left = c(1.55, 2.55),
  label = c("p=0.013 *", "p=0.851"),
  y.position = c(62, 56)  
)

inflo.height.plot<- ggplot(inflo.data, aes(y=inflo_height_cm, x=species, color=treatment))+geom_boxplot()+
  theme_minimal()+
  geom_boxplot()+
       geom_point (position=position_dodge(width=0.76))+ 
  theme( panel.grid.major.x = element_blank(),  # Remove vertical grid lines
    panel.grid.minor = element_blank(),    # Remove minor grid lines
    panel.grid.major.y = element_line(color = "grey80", size = 0.5),
    legend.position = "right")+
   labs(y="Info Height (cm)", x="", color="Nitrogen") +
   scale_color_manual(name="Nitrogen Level", 
          labels= c( HIGH= "water + nitrogen", LOW= "water only"), 
          values=c(HIGH="#E69F00", LOW="#0072B2")) +
  geom_bracket(data = brackets.inflo, aes(xmin = right, xmax = left, label = label, y.position = y.position),
               size = 0.5, bracket.shorten = 0.6, label.size = 4, inherit.aes = FALSE) +
   scale_y_continuous(limits = c(0, NA), expand=expansion(mult=c(0.02,0.2))) + 
  scale_x_discrete(
    labels = c(
      "TEN" = expression(italic("I. tenuituba")),
      "AGG" = expression(italic("I. aggregata"))
    )) +  
  theme(
    axis.text.x = element_text(size = 11), # bigger x-axis text
    legend.text = element_text(size = 10) 
  )
  
print(inflo.height.plot)
ggsave(inflo.height.plot, file="plots-nitrogen/NEXP1.greenhouse.inflo.png",  width=120, height=110, units= 'mm', dpi=600)
```

##Total Flower number
```{r}
total.flower<-read_sheet("1nyUqE-4dVr-81-MFMuAi5-dU4BIYZfuzF8xDEoFyJIk", sheet="Flower Number") %>% 
   left_join(exp1.treatments.greenhouse) %>% 
  #set treatment levels
   mutate(treatment=factor(treatment, levels=c("LOW","HIGH"))) %>% 
   mutate(species=dplyr::recode(species, "agg"="AGG", "ten"= "TEN")) %>% 
  #create column with speciestreatment aka S.T
   mutate(S.T=paste0(species,treatment)) 


flower.num.lm=lm(flower_num~treatment*species, data=total.flower)
  
Anova(flower.num.lm, type=3)#Type 3

#....................................................

flower.model <- total.flower %>%
  mutate(S.T = factor(S.T)) %>%  # Ensure S.T (species:treatment combo) is a factor
  pivot_longer(flower_num, names_to = "trait") %>%  # Pivot to long format
  group_by(trait) %>% 
  nest() %>%  # Nest data by trait to apply models separately
  mutate(
    # Fit a linear model of flower trait ~ species:treatment
    flower.lm = map(data, ~lm(value ~ S.T, data = .x))) %>% 

    # Get Type III ANOVA table from the linear model
    mutate(flower.Anova = map(flower.lm, ~tidy(Anova(.x, type = 3)))) %>% 

    # Run specific pairwise contrasts using glht with FDR correction
  mutate(flower.glht= map(flower.lm, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c("AGGLOW - AGGHIGH == 0",
                                             "TENLOW - TENHIGH ==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
     # Get estimated marginal means for each species:treatment group    
   mutate(flower.emmeans=map(flower.lm, ~tidy(emmeans(.x, "S.T")))) #gives means so we can do % change from high to low



# Format glht table for display: drop unnecessary columns, unnest glht results, and mark significance
 
flower.glht<- flower.model %>% dplyr::select(-data,-flower.lm, -flower.Anova, -flower.emmeans) %>% 
      unnest(flower.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )  

# Display glht contrast results in a table
kable(flower.glht, digits=3) #spits out model into table


# Format and display emmeans results (estimated marginal means)
flower.number.emmeans<-flower.model %>% dplyr::select(-data,-flower.lm, -flower.Anova, -flower.glht) %>% 
      unnest(flower.emmeans)  

kable(flower.number.emmeans, digits = 3)

# From the emmeans table, you can now manually calculate % change: (HIGH / LOW) - 1
# For example: AGG HIGH / LOW = 48.66 / 40 = 1.2165 → 21.65% increase


#...................................................


#plot
#facet.labs <- c(AGG="I. aggregata", TEN="I. tenuituba")

#add brackets to plot
brackets.flowernum <- tibble(
  right = c(0.45, 1.45), 
  left = c(1.55, 2.55),
  label = c("p=0.0285 *", "p=0.53"),
  y.position = c(350, 420)  
)

flower.number.plot<- ggplot(total.flower, aes(y=flower_num, x=species, color=treatment))+geom_boxplot()+ 
    geom_point (position=position_dodge(width=0.76))+ 
    scale_x_discrete(
    labels = c(
      "TEN" = expression(italic("I. tenuituba")),
      "AGG" = expression(italic("I. aggregata"))
    )) +
 scale_color_manual(name="Nitrogen Level", 
          labels= c( HIGH= "water + nitrogen", LOW= "water only"), 
          values=c(HIGH="#E69F00", LOW="#0072B2"))+
  theme_minimal()+
  theme( panel.grid.major.x = element_blank(),  # Remove vertical grid lines
    panel.grid.minor = element_blank(),    # Remove minor grid lines
    panel.grid.major.y = element_line(color = "grey80", size = 0.5),
    legend.position = "right")+
   labs(y="Flower Number", x="", color="Nitrogen") +
  geom_bracket(data = brackets.flowernum, aes(xmin = right, xmax = left, label = label, y.position = y.position),
               size = 0.5, bracket.shorten = 0.6, label.size = 4, inherit.aes = FALSE) +
    scale_y_continuous(limits = c(0, NA), expand=expansion(mult=c(0.02,0.2))) +
  theme(
  axis.text.x = element_text(size = 10)  # <- Increase or decrease this number
)
  

print(flower.number.plot)

ggsave(flower.number.plot, file="plots-nitrogen/NEXP1.greenhouse.flowernumber.png",  width=120, height=110, units= 'mm', dpi=600)

#nitrogen only p= 0.0074

#######old way#############################
#make data frame with ls means
#flower.num.N= as.data.frame(lsmeans(flower.nub.lmer, pairwise~ nitrogen|microbe, adjustSigma = TRUE, adjust = "tukey")$lsmeans)

#contrast of treatment (high vs low N) in each species (TEN vs AGG), groups a, b
#multcomp::cld(lsmeans(flower.nub.lmer, pairwise~ nitrogen|microbe, adjustSigma = TRUE, adjust = "tukey")$lsmeans ,
#alpha=0.05, Letters=letters,adjust="tukey")

###### glht #########

#library (tibble)

#need treatmentspecies so paste0 
flowerdatamutate <- as_tibble(total.flower)%>%mutate(treatmentspecies=factor(paste0(treatment,species))) #paste0 because glht doesn't like spaces in data sets

#model to put into glht bc glht doesn't like spaces
flower.nub.lm.glht=lm(flower_num~treatmentspecies, data=flowerdatamutate)

summary(multcomp::glht(flower.nub.lm.glht, linfct=multcomp::mcp(treatmentspecies="Tukey")), test=multcomp::adjusted(type="fdr"))


#Fig.flower.number
#ggsave(Fig.flower.number, file='plot_flower_number.jpg', width=110, height=110, units= 'mm', dpi=300)


```



#EFFECT SIZE  
effect size - log2 fold change

## mega emmeans
```{r}

library(dplyr)
library(tibble)
library(purrr)


mega.emmeans.long <- bind_rows(
  inflo.emmeans,
  flower.number.emmeans,
  nectar.emmeans,
  nectar.emmeans.field,
  morph.emmeans,
  morph.emmeans.field
) %>%
  # Create species column & rename lowercase to uppercase
  mutate(
    species = case_when(
      str_detect(S.T, "^AGG") ~ "AGG",
      str_detect(S.T, "^TEN") ~ "TEN",
      str_detect(S.T, "^Ihyb") ~ "Ihyb",
      str_detect(S.T, "^Lagg") ~ "Lagg",
      TRUE ~ NA_character_
    ),
    treatment = case_when(
      str_detect(S.T, "HIGH$") ~ "nitrogen",
      str_detect(S.T, "LOW$") ~ "watercontrol",
      str_detect(S.T, "ambient$") ~ "ambient",
      str_detect(S.T, "nitrogen$") ~ "nitrogen",
      str_detect(S.T, "watercontrol$") ~ "watercontrol",
      TRUE ~ NA_character_
    ),
    experiment = case_when(
      species %in% c("AGG", "TEN") ~ "greenhouse",
      species %in% c("Ihyb", "Lagg") ~ "field",
      TRUE ~ NA_character_  # catch anything unexpected
    )
  )

#calculate log2 fold change

log2_fc_table <- mega.emmeans.long %>%
  dplyr::select(trait, treatment, estimate, experiment, species, treatment) %>%
  pivot_wider(
    names_from = treatment,
    values_from = estimate,
    values_fn = mean
  ) %>%
  mutate(log2_N_WC = log2(nitrogen / watercontrol)) %>% 
  mutate(log2_WC_AM = log2(watercontrol / ambient))

#download dataframe to save as table in google sheets
#write.csv(log2_fc_table, "log2_fc_table.csv", row.names = FALSE)

#long version for ggplot formatting
log2_fc_table_long <- log2_fc_table %>%
  pivot_longer(cols = c(log2_N_WC, log2_WC_AM), 
               names_to = "effect_size_type", 
               values_to = "effect_size") %>% 
  dplyr::select(trait, species, effect_size_type, effect_size, experiment) 

#PLOT
# reorder species for plot
log2_fc_table_long$species <- factor(log2_fc_table_long$species, 
                                     levels = c("AGG", "TEN", "Lagg",  "Ihyb"))
# Reorder traits manually for plot
trait_order <- c("corolla_width", "corolla_length", "max_anther", "min_anther", 
                 "sepal_width", "style_length", "nectar.production", 
                 "concentration", "inflo_height_cm", "flower_num")

# Reorder the traits in the data
log2_fc_table_long$trait <- factor(log2_fc_table_long$trait, levels = trait_order)

#COMBINED 
#combined watercontrol + ambient for field 

#species as columns
ggplot(log2_fc_table_long, aes(x = effect_size, y = fct_rev(trait), fill = species)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Bars instead of points
  geom_vline(aes(xintercept = 0, linetype = "solid"), show.legend = FALSE) +  # Dashed line for each species
  labs(x = "Log2 Fold Change Effect Size", y = "Trait", 
       title = "Trait-level response to S.T treatments") +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(cols = vars(species)) + # Species as facet, on top
    theme(
    legend.text = element_text(size = 14),    # Increase legend text size
    legend.title = element_text(size = 16),   # Increase legend title size
    axis.text = element_text(size = 12),      # Increase general axis text size
    axis.title = element_text(size = 14),     # Increase axis title size
    strip.text = element_text(size = 16),     # Increase facet label text size
    axis.text.x = element_text(size = 16),    # Increase x-axis text size
    axis.text.y = element_text(size = 16)     # Increase y-axis text size
  )

#species as rows
ggplot(log2_fc_table_long, aes(x = effect_size, y = fct_rev(trait), fill = species)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Bars instead of points
  geom_vline(aes(xintercept = 0, linetype = "solid"), show.legend = FALSE) +  # Dashed line for each species
  labs(x = "Log2 Fold Change Effect Size", y = "Trait") +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(rows = vars(species)) + # Species as rows on right
   theme(
    legend.text = element_text(size = 14),    # Increase legend text size
    legend.title = element_text(size = 16),   # Increase legend title size
    axis.text = element_text(size = 12),      # Increase axis text size
    axis.title = element_text(size = 14),     # Increase axis title size
    strip.text = element_text(size = 16)      # Increase facet label text size
  )

#SEPARATE FACETS- log2(Nitrogen / watercontrol) & log2(watercontrol / ambient)
ggplot(log2_fc_table_long, aes(x = effect_size, y = fct_rev(trait), fill = species)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Bars instead of points
  geom_vline(aes(xintercept = 0, linetype = "solid"), show.legend = FALSE) +  # Dashed line for each species
  labs(x = "Log2 Fold Change Effect Size", y = "Trait") +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(cols = vars(effect_size_type), rows = vars(species))

```

##GH morph -log2
```{r}

morph.models.effects <- morphdata %>%
  pivot_longer(all_of(all.morph.traits), names_to = "trait") %>%
  nest_by(trait) %>%
  mutate(
    lmer_model = list(lmer(value ~ S.T + (1 | plantid), data = data)),
    emmeans_tbl = list(emmeans(lmer_model, ~ S.T) %>% as.data.frame())
  ) %>%
  unnest(emmeans_tbl) %>%
  select(trait, S.T, emmean = emmean)

#TODO: either make mega data set bindrows() OR make bindrows for emmans of all traits

#calculate log2 fold change

log2_effect_size_summary <- morph.models.effects %>%
  pivot_wider(names_from = S.T, values_from = emmean) %>% # Pivot wide for effect size calc
  mutate(
    AGG_effect_size_log2 = log2(AGGHIGH / AGGLOW),
    TEN_effect_size_log2 = log2(TENHIGH / TENLOW) 
  ) %>%
  select(trait, AGG_effect_size_log2, TEN_effect_size_log2)

#Pivot longer for ggplot
log2_effect_sizes_long <- log2_effect_size_summary %>%
  pivot_longer(cols = c(AGG_effect_size_log2, TEN_effect_size_log2),
               names_to = "treatment", values_to = "effect_size") %>%
  mutate(treatment = str_replace(treatment, "_effect_size", "")) %>% 
      drop_na(effect_size)

#plot
ggplot(log2_effect_sizes_long, aes(x = effect_size, y = fct_rev(trait), fill = treatment)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Effect size Log2 Fold Change", y = NULL,
       title = "Trait-level response to S.T treatments") +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2")


```

##field morph -log2

#% CHANGE- emmeans

```{r}

percent.change <- log2_fc_table %>%  
    mutate(
    # Create the new column for nitrogen - watercontrol
    N_WC_percent_change = ((nitrogen / watercontrol)-1)*100,  
    
    # Create the new column for watercontrol - ambient
    WC_AM_percent_change = (watercontrol / ambient)
  ) 

```

# ____________

#... extra ........

#Field- Morph models
##Field - Model for All the traits - ONLY nitrogen, ambient = water effect
```{r}
library(broom)
all.morph.traits.field <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width")

morph.models.field <- morphdata.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% group_by(trait)          %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~S.T2 + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht= map(morph.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T2= 
                                           c("Laggambient - Laggnitrogen == 0",
                                             "Ihybambient - Ihybnitrogen == 0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      #formatting for kable below
      dplyr::select(-data,-morph.trait.lmer, -morph.Anova) %>% 
      unnest(morph.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )

kable(morph.models.field, digits=3) #spits out model into table

#boxplot for averages of all traits
allmorphtraits.field <- morphdata1.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color= str_to_title(treatment), x=str_to_title(site), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width", inflo_height_cm="Inforecence Height (cm)")))+
    labs(y="Length (mm)", x="Nitrogen", color="Nitrogen") +
  scale_color_manual(name="Nitrogen Level", 
          labels= c(ambient="No water", nitrogen= "water + nitrogen", watercontrol= "water only"), 
          values=c(ambient="#457A26",nitrogen="#E69F00", watercontrol="#0072B2"))+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80", size = 0.5),
        legend.position = "bottom")
  #scale_y_continuous(limits=c(0, NA))

print(allmorphtraits.field)

ggsave(allmorphtraits.field, file="plots-nitrogen/NEXP1.field.allmorphtraits.png",  width=150, height=140, units= 'mm', dpi=600)

#difference in corolla width is due to watering plant not due to nitrogen 

```

##Field - Model for All the traits - ONLY water =(nitrogen+water control), ambient 
```{r}
library(broom)
all.morph.traits.field <- c("corolla_length", "corolla_width", "style_length", "max_anther", "min_anther","sepal_width")

morph.models.field <- morphdata.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% group_by(trait)          %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~S.T3 + (1|plantid), data=.x))) %>% 
      mutate(morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(morph.glht= map(morph.trait.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T3= 
                                           c("Laggambient - Laggwater == 0",
                                             "Ihybambient - Ihybwater == 0" 
                                              ))), test=multcomp::adjusted(type="fdr"))))) %>% 
      #formatting for kable below
      dplyr::select(-data,-morph.trait.lmer, -morph.Anova) %>% 
      unnest(morph.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )

kable(morph.models.field, digits=3) #spits out model into table

#boxplot for averages of all traits
allmorphtraits.field <- morphdata1.field %>% pivot_longer(all_of(all.morph.traits.field), names_to = "trait") %>% 
    #filter(trait%in% c("corolla_length", "max_anther")) %>% 
    ggplot(aes(color= str_to_title(treatment), x=str_to_title(site), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(corolla_length="Corolla Length", corolla_width= "Corolla Width", style_length="Style Length", max_anther="Max Stamen", min_anther= "Min Stamen",sepal_width="Sepal Width", inflo_height_cm="Inforecence Height (cm)")))+
    labs(y="Length (mm)", x="Nitrogen", color="Nitrogen") +
  scale_color_manual(values = c("orange","purple","blue"))+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), 
        panel.grid.major.y = element_line(color = "grey80", size = 0.5),
        legend.position = "bottom")
  #scale_y_continuous(limits=c(0, NA))

print(allmorphtraits.field)

ggsave(allmorphtraits.field, file="plots-nitrogen/NEXP1.field.allmorphtraits.png",  width=150, height=140, units= 'mm', dpi=600)

#difference in corolla width is due to watering plant not due to nitrogen 

```

#Cohens D Effect Size (mean high-low/SD)
##GH effect size all morph
```{r}
# Step 0: Get SDs per trait from raw data
trait_sd_df <- morphdata %>%
  pivot_longer(all_of(all.morph.traits), names_to = "trait") %>%
  group_by(trait,species) %>%
  summarise(trait_sd = sd(value, na.rm = TRUE))

morph.models.effects <- morphdata %>%
  pivot_longer(all_of(all.morph.traits), names_to = "trait") %>%
  nest_by(trait) %>%
  mutate(
    lmer_model = list(lmer(value ~ S.T + (1 | plantid), data = data)),
    emmeans_tbl = list(emmeans(lmer_model, ~ S.T) %>% as.data.frame())
  ) %>%
  unnest(emmeans_tbl) %>%
  select(trait, S.T, emmean = emmean)

# Join in SDs
morph.models.effects <- morph.models.effects %>%
  mutate(species= str_sub(S.T, 1,3)) %>% 
  left_join(trait_sd_df, by = c("trait","species"))

# Pivot wide for effect size calc
effect_size_summary <- morph.models.effects %>%
  pivot_wider(names_from = S.T, values_from = emmean) %>%
  mutate(
    AGG_effect_size = (AGGHIGH - AGGLOW) / trait_sd,
    TEN_effect_size = (TENHIGH - TENLOW) / trait_sd
  ) %>%
  select(trait, AGG_effect_size, TEN_effect_size)


effect_sizes_long <- effect_size_summary %>%
  pivot_longer(cols = c(AGG_effect_size, TEN_effect_size),
               names_to = "treatment", values_to = "effect_size") %>%
  mutate(treatment = str_replace(treatment, "_effect_size", "")) %>% 
      drop_na(effect_size)

#plot
ggplot(effect_sizes_long, aes(x = effect_size, y = fct_rev(trait), fill = treatment)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Effect size (standardized mean difference)", y = NULL,
       title = "Trait-level response to S.T treatments") +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2")


```


##GH corolla width emmeans & effect size
```{r}
#set up for emmeans
#run model with only corolla width

morph.models.corolla.width <- morphdata %>% pivot_longer(all_of(all.morph.traits), names_to = "trait") %>% 
      filter(trait%in% c("corolla_width")) %>% 
      nest() %>% 
      mutate(corolla.lmer=map(data, ~lmer(value~S.T + (1|plantid), data=.x))) %>% 
      mutate(corolla.Anova= map(corolla.lmer, ~tidy(Anova(.x, type=3)))) %>%  
      mutate(corolla.glht= map(corolla.lmer, ~tidy(multcomp:::summary.glht(multcomp::glht(.x, linfct=multcomp::mcp(S.T= 
                                           c("AGGLOW - AGGHIGH == 0",
                                             "TENLOW - TENHIGH ==0"))), test=multcomp::adjusted(type="fdr"))))) %>% 
      mutate(emmeans=map(corolla.lmer, ~tidy(emmeans(.x, "S.T")))) #gives means so we can do % change from high to low

      #formatting for kable below
corolla.width.model <-morph.models.corolla.width  %>% dplyr::select(-data,-corolla.lmer, -corolla.Anova, -emmeans) %>% 
      unnest(corolla.glht) %>% 
      mutate(significant = ifelse(adj.p.value<0.05, "*","") )  
 
#glht contrasts     
kable(corolla.width.model, digits=3) #spits out model into table

#emmeans table 
corolla.width.emmeans<-morph.models.corolla.width %>% dplyr::select(-data,-corolla.lmer, -corolla.Anova, -corolla.glht) %>% 
      unnest(emmeans)  
    
kable(corolla.width.emmeans, digits=3) #spits out model into table
#now divide 'estimate' high by low to get % change
#AGG high/low = 48.66/40 = 1.2165 --> 21.65% increase in high????

#Effect Size
# Get SD of raw corolla_width
corolla.sd <- morphdata %>%
  filter(!is.na(corolla_width)) %>%
  summarize(sd = sd(corolla_width)) %>%
  pull(sd)

# Extract emmeans and calculate effect sizes
effect_sizes <- corolla.width.emmeans %>%
  select(S.T, estimate) %>%
  pivot_wider(names_from = S.T, values_from = estimate) %>%
  mutate(
    AGG_effect_size = (AGGHIGH - AGGLOW) / corolla.sd,
    TEN_effect_size = (TENHIGH - TENLOW) / corolla.sd
  )

# View result
kable(effect_sizes, digits = 3)


```
